---
title: "Load and Combine CSV Files"
author: "Callum Weinberg"
date: "January 30, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# This is necessary given the nested directories.
# Enter your working directory here
knitr::opts_knit$set(root.dir = "/home/clw/Documents/UCSB/Quarter6/PSTAT 231/Final Project/PSTAT231_Final")
```

# README

The initial version of this cleaning will just keep the variables that will be considered for the analysis. A future version of this file may fully load all the columns from the data source.

# Libraries

```{r libraries}
library(plyr)
library(dplyr)
library(tidyr)
library(assertr)
```

# Functions

```{r functions}
# From https://cran.r-project.org/web/packages/assertr/vignettes/assertr.html
not.missing.p <- function(x) if(is.na(x)) return(FALSE)
```

# Manual Crosswalks

```{r load_manual_crosswalks}
## Energy and Minerals Crosswalk
energy_minerals_crosswalk_manual = read.csv("Crosswalks/Energy_Minerals_Crosswalk.csv")
energy_minerals_crosswalk_mineral_manual = read.csv("Crosswalks/Energy_Minerals_Crosswalk_mineral_dataset.csv")

## Land and Agriculture Crosswalks
land_agriculture_preliminary_manual = read.csv("Crosswalks/land_agricultural_preliminary.csv")
land_agriculture_final_manual = read.csv("Crosswalks/land_agricultural_final.csv")

## Final Crosswalk for Mapping Categories
## Includes World Bank Data
final_crosswalk = read.csv("Crosswalks/02_final_crosswalk.csv")
final_crosswalk = final_crosswalk %>%
  rename(Country = Country_Mapping)
# Get version with only Country and ID
final_crosswalk_Country_Only = final_crosswalk %>%
  select(Country, Country_ID_Final)
```



# Air and Climate Data

## Load and Clean Air and Climate Data

```{r load_air_and_climate}
## CH_4 Emissions
CH4_emissions = read.csv("Raw_Data/Air and Climate/CH4_Emissions.csv")

CH4_emissions = CH4_emissions %>%
  rename(Country_ID = Country.ID,
         CH4_latest_year = X.28, 
         CH4 = CH4.emissions..latest.year, 
         CH4_change_1990 = X..change.since.1990, 
         CH4_per_capita = CH4.emissions..per.capita...latest.year) %>%
  select(Country_ID,Country,CH4_latest_year,CH4,CH4_change_1990,CH4_per_capita) %>%
  slice(2:n()) %>%
  assert(not.missing.p, Country_ID)

## CO2 Emissions
CO2_emissions = read.csv("Raw_Data/Air and Climate/CO2_Emissions.csv")

CO2_emissions = CO2_emissions %>%
  rename(Country_ID = Country.ID,
         CO2_latest_year = X.28, 
         CO2 = CO2.emissions..latest.year, 
         CO2_change_1990 = X..change.since.1990, 
         CO2_per_capita = CO2.emissions..per.capita...latest.year) %>%
  select(Country_ID,Country,CO2_latest_year,CO2,CO2_change_1990,CO2_per_capita) %>%
  slice(2:n()) %>%
  assert(not.missing.p, Country_ID)

## GHG Emissions
GHG_emissions = read.csv("Raw_Data/Air and Climate/GHG_Emissions.csv")

GHG_emissions = GHG_emissions %>%
  rename(Country_ID = Country.ID,
         GHG_latest_year = X.28, 
         GHG = GHG.total.without.LULUCF..latest.year, 
         GHG_change_1990 = X..change.since.1990, 
         GHG_per_capita = GHG.emissions.per.capita...latest.year) %>%
  select(Country_ID,Country,GHG_latest_year,GHG,GHG_change_1990,GHG_per_capita) %>%
  slice(2:n()) %>%
  assert(not.missing.p, Country_ID)

## GHG Emissions
GHG_emissions = read.csv("Raw_Data/Air and Climate/GHG_Emissions.csv")
# GHG_sector_total should be the same as GHG, if not investigate

GHG_emissions = GHG_emissions %>%
  rename(Country_ID = Country.ID,
         GHG_latest_year = X.28, 
         GHG = GHG.total.without.LULUCF..latest.year,
         GHG_change_1990 = X..change.since.1990, 
         GHG_per_capita = GHG.emissions.per.capita...latest.year) %>%
  select(Country_ID,Country,GHG_latest_year,GHG,GHG_change_1990,GHG_per_capita) %>%
  slice(2:n()) %>%
  assert(not.missing.p, Country_ID)

## GHG Emissions by Sector
# Add this in
# GHG_emissions_by_sector = read.csv("Data/Air and Climate/GHG_Emissions_by_Sector.csv")
# GHG_sector_total should be the same as GHG, if not investigate

## N2O Emissions
N2O_emissions = read.csv("Raw_Data/Air and Climate/N2O_Emissions.csv")

N2O_emissions = N2O_emissions %>%
  rename(Country_ID = Country.ID,
         N2O_latest_year = X.28, 
         N2O = N2O.emissions..latest.year, 
         N2O_change_1990 = X..change.since.1990, 
         N2O_per_capita = N2O.emissions..per.capita...latest.year) %>%
  select(Country_ID,Country,N2O_latest_year,N2O,N2O_change_1990,N2O_per_capita) %>%
  slice(2:n()) %>%
  assert(not.missing.p, Country_ID)

## NOx Emissions
# There are some blank lines in this
# file that need to be skipped
NOx_emissions = read.csv("Raw_Data/Air and Climate/NOx_Emissions.csv")

NOx_emissions = NOx_emissions %>%
  rename(Country_ID = Country.ID,
         NOx_latest_year = X.28, 
         NOx = NOx.emissions..latest.year, 
         NOx_change_1990 = X..change.since.1990, 
         NOx_per_capita = NOx..emissions.per.capita...latest.year) %>%
  select(Country_ID,Country,NOx_latest_year,NOx,NOx_change_1990,NOx_per_capita) %>%
  slice(2:173) %>%
  assert(not.missing.p, Country_ID)

## SO2 Emissions
SO2_emissions = read.csv("Raw_Data/Air and Climate/SO2_emissions.csv")

SO2_emissions = SO2_emissions %>%
  rename(Country_ID = Country.ID,
         SO2_latest_year = X.28, 
         SO2 = SO2.emissions..latest.year, 
         SO2_change_1990 = X..change.since.1990, 
         SO2_per_capita = SO2.emissions.per.capita..latest.year) %>%
  select(Country_ID,Country,SO2_latest_year,SO2,SO2_change_1990,SO2_per_capita) %>%
  slice(2:143) %>%
  assert(not.missing.p, Country_ID)
```

## Air and Climate Crosswalk

```{r crosswalk_air_and_climate}
# Append all of the Datasets
crosswalk_air_climate = rbind.fill(CH4_emissions,
                                   CO2_emissions,
                                   GHG_emissions,
                                   N2O_emissions,
                                   NOx_emissions,
                                   SO2_emissions)

# Create the Crosswalk
crosswalk_air_climate = crosswalk_air_climate %>%
  select(Country_ID, Country) %>%
  distinct()

# Check if there are any Duplicates 
# This would mean that for the air and climate datasets
# there are either repeated countries, IDs, or that
# ID is not unique to country between the datasets
# and Vice versa
dim(crosswalk_air_climate[duplicated(crosswalk_air_climate$Country_ID),])[1] == 0
dim(crosswalk_air_climate[duplicated(crosswalk_air_climate$Country),])[1] == 0
```

## Combine Air and Climate

```{r combine_air_and_climate}
# Start with the Above Created Crosswalk
# And Merge on Each of the Datasets
combined_air_climate = 
  left_join(crosswalk_air_climate,CH4_emissions, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_air_climate = 
  left_join(combined_air_climate,CO2_emissions, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_air_climate = 
  left_join(combined_air_climate,GHG_emissions, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_air_climate = 
  left_join(combined_air_climate,N2O_emissions, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_air_climate = 
  left_join(combined_air_climate,NOx_emissions, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_air_climate = 
  left_join(combined_air_climate,SO2_emissions, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)

# Export Data
save(combined_air_climate, file="Intermediate_Data/001_combined_air_climate.Rdata")
write.csv(combined_air_climate,"Intermediate_Data/001_combined_air_climate.csv", row.names = FALSE)

# Merge in Final Crosswalk to Get Universal ID
combined_air_climate_clean = 
  left_join(final_crosswalk_Country_Only,combined_air_climate, by = "Country") %>%
  filter(!is.na(Country_ID)) %>%
  select(-Country, -Country_ID) %>%
  filter(Country_ID_Final != 999) #Excluded before merging

# Clean Up Environment
remove(crosswalk_air_climate,combined_air_climate, CH4_emissions, CO2_emissions,
       GHG_emissions,N2O_emissions, NOx_emissions, SO2_emissions)
```



# Energy and Minerals

## Load and Clean Energy and Minerals Data

```{r load_energy}
## Renewable Percentage
# Current problem is to model 2017 renewable energy
# so only keeping 2017 variable in this iteration

# Also there are some empty rows and the end of the
# data set
renewable_percentage = read.csv("Raw_Data/Energy and Minerals/Renewable elec production percentage.csv")

renewable_percentage = renewable_percentage %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area, 
         renewable_percent_2017 = X2017) %>%
  select(Country_ID,renewable_percent_2017) %>%
  slice(1:224) %>%
  assert(not.missing.p, Country_ID) %>%
  left_join(energy_minerals_crosswalk_manual, by = "Country_ID")

## Energy Indicators
# YEAR NOT INDICATED
energy_indicators = read.csv("Raw_Data/Energy and Minerals/Energy Indicators.csv")

energy_indicators = energy_indicators %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area, 
         energy_supply_petajoules = Energy.supply..petajoules.,
         energy_supply_per_capita_gigajoules = Energy.supply..per.capita...gigajoules.per.capita.,
         contribution_renewable_to_electric_production = Conribution.of.renewables.to.electricity.production....) %>%
  select(Country_ID,energy_supply_petajoules,
         energy_supply_per_capita_gigajoules, 
         contribution_renewable_to_electric_production) %>%
  slice(1:217) %>%
  assert(not.missing.p, Country_ID) %>%
  left_join(energy_minerals_crosswalk_manual, by = "Country_ID")

## Energy Intensity
energy_intensity = read.csv("Raw_Data/Energy and Minerals/Energy intensity measured in terms of primary energy and GDP.csv")

energy_intensity = energy_intensity %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area, 
         energy_intensity_2017 = X2017) %>%
  select(Country_ID, energy_intensity_2017) %>%
  assert(not.missing.p, Country_ID) %>%
  left_join(energy_minerals_crosswalk_manual, by = "Country_ID")

## Energy Supply Per Capita
energy_supply_per_capita = read.csv("Raw_Data/Energy and Minerals/Energy supply per capita.csv")

energy_supply_per_capita = energy_supply_per_capita %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area, 
         energy_supply_per_capita_2017 = X2017) %>%
  select(Country_ID,energy_supply_per_capita_2017) %>%
  slice(1:222) %>%
  assert(not.missing.p, Country_ID) %>%
  left_join(energy_minerals_crosswalk_manual, by = "Country_ID")

## Energy Supply
energy_supply = read.csv("Raw_Data/Energy and Minerals/Energy supply.csv")

energy_supply = energy_supply %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area, 
         energy_supply_2017 = X2017) %>%
  select(Country_ID,energy_supply_2017) %>%
  slice(1:222) %>%
  assert(not.missing.p, Country_ID) %>%
  left_join(energy_minerals_crosswalk_manual, by = "Country_ID")

## Contribution of Mining to Value Added
# Only keeping 2017
# Missing ID: will need to fix this
mining_percentage = read.csv("Raw_Data/Energy and Minerals/Contribution of mining to value added.csv")

mining_percentage = mining_percentage %>%
  rename(Country = Country.and.area, 
         mining_value_2017 = X2017) %>%
  select(Country,mining_value_2017) %>%
  assert(not.missing.p, Country) %>%
  left_join(energy_minerals_crosswalk_mineral_manual, 
            by = c("Country" = "Country_Corrected")) %>% # Correct Missing ID
  select(-Country) %>%
  left_join(energy_minerals_crosswalk_manual, by = "Country_ID")
```

## Energy and Minerals Crosswalk

```{r crosswalk_air_and_climate}
# Append all of the Datasets
crosswalk_energy_minerals = rbind.fill(renewable_percentage,
                                   energy_indicators,
                                   energy_intensity,
                                   energy_supply_per_capita,
                                   energy_supply, 
                                   mining_percentage)

# Create the Crosswalk
crosswalk_energy_minerals = crosswalk_energy_minerals %>%
  select(Country_ID, Country) %>%
  distinct()

#write.csv(crosswalk_energy_minerals,"Crosswalks/Base_Energy_Minerals_Crosswalk.csv", row.names = FALSE)

# Check if there are any Duplicates 
# This would mean that for the energy and minerals datasets
# there are either repeated countries, IDs, or that
# ID is not unique to country between the datasets
# and Vice versa
dim(crosswalk_energy_minerals[duplicated(crosswalk_energy_minerals$Country_ID),])[1] == 0
dim(crosswalk_energy_minerals[duplicated(crosswalk_energy_minerals$Country_Corrected),])[1] == 0
```

## Combine Energy and Minerals

```{r combine_air_and_climate}
# Start with the Above Created Crosswalk
# And Merge on Each of the Datasets
combined_energy_minerals = 
  left_join(crosswalk_energy_minerals,renewable_percentage, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_energy_minerals = 
  left_join(combined_energy_minerals,energy_indicators, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_energy_minerals = 
  left_join(combined_energy_minerals,energy_intensity, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_energy_minerals = 
  left_join(combined_energy_minerals,energy_supply_per_capita, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_energy_minerals = 
  left_join(combined_energy_minerals,energy_supply, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_energy_minerals = 
  left_join(combined_energy_minerals,mining_percentage, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)

# Export Data
save(combined_energy_minerals, file="Intermediate_Data/002_combined_energy_minerals.Rdata")
write.csv(combined_energy_minerals,"Intermediate_Data/002_combined_energy_minerals.csv", row.names = FALSE)

# Merge in Final Crosswalk to Get Universal ID
combined_energy_minerals_clean = 
  left_join(final_crosswalk_Country_Only,combined_energy_minerals, by = "Country") %>%
  filter(!is.na(Country_ID)) %>%
  select(-Country, -Country_ID) %>%
  filter(Country_ID_Final != 999) #Excluded before merging

# Clean Up Environment
remove(crosswalk_energy_minerals,combined_energy_minerals,renewable_percentage, energy_indicators,
       energy_intensity,energy_supply_per_capita,energy_supply, mining_percentage,energy_minerals_crosswalk_manual, 
       energy_minerals_crosswalk_mineral_manual)
```


# Land and Agriculture Data

## Load and Clean Land and Agriculture Data

```{r load_land_agriculture}
## Agricultural Land
agriculture_land = read.csv("Raw_Data/Land and Agriculture/Agricultural Land.csv")

# not including change in cultural land since 1990
# missing a good amount. Also not including
# agricultural land actually irrigated, also 
# missing a lot
agriculture_land = agriculture_land %>%
  rename(Country = Country,
         agricultural_area_2013 = Agricultural.area.in.2013..km2., 
         percent_land_agricultural_2013 = X..of.total.land.area.covered.by.agricultural.area.in.2013, 
         arable_land_2013 = Arable.land.in.2013..km2., 
         permanent_crops_2013 = Permanent.crops.in.2013..km2.,
         permanent_meadows_pastures_2013 = Permanent.meadows.and.pastures.in.2013..km2.) %>%
  select(Country,agricultural_area_2013,percent_land_agricultural_2013,arable_land_2013,
         permanent_crops_2013,permanent_meadows_pastures_2013) %>%
  assert(not.missing.p, Country) %>% # ID missing 
  left_join(land_agriculture_preliminary_manual, by = "Country") %>%
  assert(not.missing.p, Country_ID) %>%
  select(-Country) %>%
  left_join(land_agriculture_final_manual, by = "Country_ID")


## Consumption of Fertilizers Per Unit of 
## Agricultural Land
consumption_fertilizer = read.csv("Raw_Data/Land and Agriculture/Consumption of fertilizers per unit of agricultural land area.csv")

consumption_fertilizer = consumption_fertilizer %>%
  rename(Country = Country,
         nitrogen_consumption = Nitrogen, 
         phosphate_consumption = Phosphate, 
         potash_consumption = Potash) %>%
  select(Country,nitrogen_consumption,phosphate_consumption,potash_consumption) %>%
  assert(not.missing.p, Country) %>% # ID missing 
  left_join(land_agriculture_preliminary_manual, by = "Country") %>%
  assert(not.missing.p, Country_ID) %>%
  select(-Country) %>%
  left_join(land_agriculture_final_manual, by = "Country_ID")

## Terrestrial Protected Areas
# All 2018
terrestrial_protected_areas = read.csv("Raw_Data/Land and Agriculture/Terrestrial protected areas.csv")
  
terrestrial_protected_areas = terrestrial_protected_areas %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area,
         terrestrial_protected_areas = Terrestrial.protected.areas) %>%
  select(Country_ID,Country,terrestrial_protected_areas) %>%
  slice(3:n()) %>%
  assert(not.missing.p, Country_ID) %>%
  select(-Country) %>%
  left_join(land_agriculture_final_manual, by = "Country_ID")
```

## Land and Agricultural Crosswalk

```{r crosswalk_land_agricultural}
# Append all of the Datasets
crosswalk_land_agricultural = rbind.fill(agriculture_land,
                                   consumption_fertilizer,
                                   terrestrial_protected_areas)

# Create the Crosswalk
crosswalk_land_agricultural = crosswalk_land_agricultural %>%
  select(Country_ID, Country) %>%
  distinct()

# Check if there are any Duplicates 
# This would mean that for the energy and minerals datasets
# there are either repeated countries, IDs, or that
# ID is not unique to country between the datasets
# and Vice versa
dim(crosswalk_land_agricultural[duplicated(crosswalk_land_agricultural$Country_ID),])[1] == 0
dim(crosswalk_land_agricultural[duplicated(crosswalk_land_agricultural$Country_Corrected),])[1] == 0
```

## Combine Land Agricultural

```{r combine_air_and_climate}
# Start with the Above Created Crosswalk
# And Merge on Each of the Datasets
combined_land_agricultural = 
  left_join(crosswalk_land_agricultural,agriculture_land, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_land_agricultural = 
  left_join(combined_land_agricultural,consumption_fertilizer, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_land_agricultural = 
  left_join(combined_land_agricultural,terrestrial_protected_areas, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)

# Export Data
save(combined_land_agricultural, file="Intermediate_Data/003_combined_land_agricultural.Rdata")
write.csv(combined_land_agricultural,"Intermediate_Data/003_combined_land_agricultural.csv", row.names = FALSE)

# Merge in Final Crosswalk to Get Universal ID
combined_land_agricultural_clean = 
  left_join(final_crosswalk_Country_Only,combined_land_agricultural, by = "Country") %>%
  filter(!is.na(Country_ID)) %>%
  select(-Country, -Country_ID) %>%
  filter(Country_ID_Final != 999) #Excluded before merging

# Clean Up Environment
remove(crosswalk_land_agricultural,combined_land_agricultural,agriculture_land, 
       consumption_fertilizer,terrestrial_protected_areas, land_agriculture_preliminary_manual,
       land_agriculture_final_manual)
```


# Natural Disasters

```{r load_natural_disasters}
## climatological disasters
climatological_disasters = read.csv("Raw_Data/Natural Disasters/Climatological disasters.csv")

climatological_disasters = climatological_disasters %>%
  rename(Country_ID = CountryID,
         Country = Countries.or.areas, 
         climate_disaster_occurence_2010_2019 = Occurrence.2010.2019, 
         climate_disaster_deaths_2010_2019 = Total.deaths.2010.2019, 
         climate_disaster_affected_2010_2019 = Persons.affected.2010.2019) %>%
  select(Country_ID,Country,climate_disaster_occurence_2010_2019,
         climate_disaster_deaths_2010_2019,
         climate_disaster_affected_2010_2019) %>%
  assert(not.missing.p, Country_ID)

## geophysical disasters
geophysical_disasters = read.csv("Raw_Data/Natural Disasters/Geophysical disasters.csv")

geophysical_disasters = geophysical_disasters %>%
  rename(Country_ID = CountryID,
         Country = Countries.or.areas, 
         geophysical_disaster_occurence_2010_2019 = Occurrence.2010.2019, 
         geophysical_disaster_deaths_2010_2019 = Total.deaths.2010.2019, 
         geophysical_disaster_affected_2010_2019 = Persons.affected.2010.2019) %>%
  select(Country_ID,Country,geophysical_disaster_occurence_2010_2019,
         geophysical_disaster_deaths_2010_2019,
         geophysical_disaster_affected_2010_2019) %>%
  assert(not.missing.p, Country_ID)


## hydrological disasters
hydrological_disasters = read.csv("Raw_Data/Natural Disasters/Hydrological disasters.csv")

hydrological_disasters = hydrological_disasters %>%
  rename(Country_ID = CountryID,
         Country = Countries.or.areas, 
         hydrological_disaster_occurence_2010_2019 = Occurrence.2010.2019, 
         hydrological_disaster_deaths_2010_2019 = Total.deaths.2010.2019, 
         hydrological_disaster_affected_2010_2019 = Persons.affected.2010.2019) %>%
  select(Country_ID,Country,hydrological_disaster_occurence_2010_2019,
         hydrological_disaster_deaths_2010_2019,
         hydrological_disaster_affected_2010_2019) %>%
  assert(not.missing.p, Country_ID)


## meteorological disasters
meteorological_disasters = read.csv("Raw_Data/Natural Disasters/Meteorological disasters.csv")

meteorological_disasters = meteorological_disasters %>%
  rename(Country_ID = CountryID,
         Country = Countries.or.areas, 
         meteorological_disaster_occurence_2010_2019 = Occurrence.2010.2019, 
         meteorological_disaster_deaths_2010_2019 = Total.deaths.2010.2019, 
         meteorological_disaster_affected_2010_2019 = Persons.affected.2010.2019) %>%
  select(Country_ID,Country,meteorological_disaster_occurence_2010_2019,
         meteorological_disaster_deaths_2010_2019,
         meteorological_disaster_affected_2010_2019) %>%
  assert(not.missing.p, Country_ID)
```


## Natural Disasters Crosswalk

```{r crosswalk_land_agricultural}
# Append all of the Datasets
crosswalk_natural_disaster = rbind.fill(climatological_disasters,
                                   geophysical_disasters,
                                   hydrological_disasters,
                                   meteorological_disasters)

# Create the Crosswalk
crosswalk_natural_disaster = crosswalk_natural_disaster %>%
  select(Country_ID, Country) %>%
  distinct()

# Check if there are any Duplicates 
# This would mean that for the energy and minerals datasets
# there are either repeated countries, IDs, or that
# ID is not unique to country between the datasets
# and Vice versa
dim(crosswalk_natural_disaster[duplicated(crosswalk_natural_disaster$Country_ID),])[1] == 0
dim(crosswalk_natural_disaster[duplicated(crosswalk_natural_disaster$Country_Corrected),])[1] == 0
```

## Combine Natural Disaster

```{r combine_natural_disaster}
# Start with the Above Created Crosswalk
# And Merge on Each of the Datasets
combined_natural_disaster = 
  left_join(crosswalk_natural_disaster,climatological_disasters, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_natural_disaster = 
  left_join(combined_natural_disaster,geophysical_disasters, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_natural_disaster = 
  left_join(combined_natural_disaster,hydrological_disasters, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_natural_disaster = 
  left_join(combined_natural_disaster,meteorological_disasters, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)

# Export Data
save(combined_natural_disaster, file="Intermediate_Data/004_combined_natural_disaster.Rdata")
write.csv(combined_natural_disaster,"Intermediate_Data/004_combined_natural_disaster.csv", row.names = FALSE)

# Merge in Final Crosswalk to Get Universal ID
combined_natural_disaster_clean = 
  left_join(final_crosswalk_Country_Only,combined_natural_disaster, by = "Country") %>%
  filter(!is.na(Country_ID)) %>%
  select(-Country, -Country_ID) %>%
  filter(Country_ID_Final != 999) #Excluded before merging

# Clean Up Environment
remove(crosswalk_natural_disaster,combined_natural_disaster,climatological_disasters, 
       geophysical_disasters,hydrological_disasters, meteorological_disasters)
```

# Other Categories

## Load and Clean Other Categories

```{r load_other}
## Biodiversity Data/Marine and Terrestrial Protected areas
# row 1 is world total
protected_areas_marine_terrestrial = read.csv("Raw_Data/Biodiversity/Terrestrial_Marine protected areas.csv")

protected_areas_marine_terrestrial = 
  protected_areas_marine_terrestrial %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area,
         protected_areas_marine_terrestrial_latest_year = latest.year.available, 
         protected_areas_marine_terrestrial = Terrestrial.and.marine.protected.areas....of.total.territorial.area.) %>%
  slice(2:n()) %>%
  assert(not.missing.p, Country_ID)

## Marine and Coastal Protected areas
# first row is world total
protected_areas_marine_coastal = read.csv("Raw_Data/Marine and Coastal Areas/Marine protected areas.csv")

protected_areas_marine_coastal = 
  protected_areas_marine_coastal %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area,
         protected_areas_marine_latest_year = latest.year.available, 
         protected_areas_marine = Marine.protected.areas....of.territorial.waters.) %>%
  slice(2:n()) %>%
  assert(not.missing.p, Country_ID)

## Forests Data
# deforestaion and fire areas not included due to 
# significant number of missing values
# row 1 is world total
forest_areas = read.csv("Raw_Data/Forests/Forest Area.csv")

forest_areas = 
  forest_areas %>%
  rename(Country_ID = CountryID,
         Country = Country.and.Area,
         forest_area_2015 = Forest.Area..2015..1000.ha., 
         forest_area_2020 = Forest.Area..2020..1000.ha.,
         forest_total_land_area_2020 = Total.Land.Area..2020..1000.ha.) %>%
  select(Country_ID,Country,forest_area_2015,forest_area_2020,forest_total_land_area_2020) %>%
  slice(2:n()) %>%
  assert(not.missing.p, Country_ID)


## Governance
governance = read.csv("Raw_Data/Governance/Governance.csv")

governance = governance %>%
  rename(Country_ID = CountryID,
         Country = Country.and.area) %>%
  rename_with(~gsub(".", "_", .x, fixed = TRUE)) %>%
  assert(not.missing.p, Country_ID)
```

## Other Crosswalk

```{r crosswalk_other}
# Append all of the Datasets
crosswalk_other = rbind.fill(protected_areas_marine_terrestrial,
                                   protected_areas_marine_coastal,
                                   forest_areas,
                                   governance)

# Create the Crosswalk
crosswalk_other = crosswalk_other %>%
  select(Country_ID, Country) %>%
  distinct()

# Check if there are any Duplicates 
# This would mean that for the air and climate datasets
# there are either repeated countries, IDs, or that
# ID is not unique to country between the datasets
# and Vice versa
dim(crosswalk_other[duplicated(crosswalk_other$Country_ID),])[1] == 0
dim(crosswalk_other[duplicated(crosswalk_other$Country),])[1] == 0
```

## Combine Others

```{r combine_air_and_climate}
# Start with the Above Created Crosswalk
# And Merge on Each of the Datasets
combined_other = 
  left_join(crosswalk_other,protected_areas_marine_terrestrial, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_other = 
  left_join(combined_other,protected_areas_marine_coastal, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_other = 
  left_join(combined_other,forest_areas, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)
combined_other = 
  left_join(combined_other,governance, by = "Country_ID") %>%
  select(-Country.y) %>%
  rename(Country = Country.x)

# Export Data
save(combined_other, file="Intermediate_Data/005_combined_other.Rdata")
write.csv(combined_other,"Intermediate_Data/005_combined_other.csv", row.names = FALSE)

# Merge in Final Crosswalk to Get Universal ID
combined_other_clean = 
  left_join(final_crosswalk_Country_Only,combined_other, by = "Country") %>%
  filter(!is.na(Country_ID)) %>%
  select(-Country, -Country_ID) %>%
  filter(Country_ID_Final != 999) #Excluded before merging

# Clean Up Environment
remove(crosswalk_other,combined_other,protected_areas_marine_terrestrial, 
       protected_areas_marine_coastal,forest_areas, governance)
```









# Combine Categories

## Create Combined Dataset

```{r combined_all}

# Create Clea Final Data and World Bank Dataset
full_data = final_crosswalk %>%
  filter(Country_ID_Final != 999) %>%
  filter(Country_Final != "Swaziland") %>%
  filter(Country_Final != "Czechia") %>%
  filter(Country_Final != "St. Helena") %>%
  filter(Country_Final != "St. Pierre-Miquelon") %>%
  select(-Country) %>%
  distinct()

# Check if there are any Duplicates 
# This would mean that for the air and climate datasets
# there are either repeated countries, IDs, or that
# ID is not unique to country between the datasets
# and Vice versa
dim(full_data[duplicated(full_data$Country_ID_Final),])[1] == 0
dim(full_data[duplicated(full_data$Country_Final),])[1] == 0

# Create Full Dataset (Excluding 999 Countries)
full_data = left_join(full_data,combined_air_climate_clean, by = "Country_ID_Final")
full_data = left_join(full_data,combined_energy_minerals_clean, by = "Country_ID_Final")
full_data = left_join(full_data,combined_land_agricultural_clean, by = "Country_ID_Final")
full_data = left_join(full_data,combined_natural_disaster_clean, by = "Country_ID_Final")
full_data = left_join(full_data,combined_other_clean, by = "Country_ID_Final")
```


# Output Data

```{r output_data}
# Save the Data for Use in the Next File
save(full_data, file="Intermediate_Data/01_full_data.Rdata")
```