---
title: "03_Missing_Data_Imputation"
author: "Callum Weinberg"
date: "February 23, 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# This is necessary given the nested directories.
# Enter your working directory here
knitr::opts_knit$set(root.dir = "/home/clw/Documents/UCSB/Quarter6/PSTAT 231/Final Project/PSTAT231_Final")
```

# README

Missing data imputation. For now assigning median value. This will not be the final approach, but doing so to move along the analysis process

# Libraries

```{r libraries}
library(plyr)
library(dplyr)
library(tidyr)
library(assertr)
library(knitr)
library(ggplot2)
library(reshape2)
```

# Functions
```{r}
# Define Median Replacement function (for a matrix)
median_replace = function(x) {
  for(i in 1:ncol(x)) {
    # Note: this will break if an entire column is missing values
    x[which(is.na(x[,i])),i] = median(x[,i],na.rm=TRUE)
  }
  return(x)
}
```

# Load the Data

```{r load_data}
## Load the Data from the 02 File
load(file="Intermediate_Data/02_limited_data.Rdata")
```

# Remove contribution_renewable_to_electric_production

```{r}
# Highly Correlated, remove from data
limited_data = limited_data[,-12]

```

# Impute Missing Data

The goal here is to predict each missing value using a linear model of all other observations, where any missings in the fit for the model are substituted with the median (but only for the fitting and prediction stage).

```{r impute_data_updated, warning = FALSE}
# Count Number of Missings
# colSums(is.na(limited_data))

# Define Model Data
model_data = limited_data

# Loop Over Each Variable Not in 
for(i in 7:ncol(limited_data)) {

  # Get Vector of Predictors (Column Position in limited_data)
  predictors = c(7:34)
  predictors = predictors[predictors != i]

  # Fit the Linear Model on All Precitors (excluding the)
  # variable that is being interpolated
  fit = 
    lm(data = limited_data, 
       as.matrix(limited_data[,i]) ~ as.matrix(median_replace(limited_data[,predictors])))
  
  # Predict
  prediction = predict(fit,as.data.frame(median_replace(limited_data[,predictors])))
  
  # No Predictions should be less than 0
  # This is based on the variables included, and
  # may change if more variables are included
  negative_indeces = which(prediction < 0)
  prediction[negative_indeces] = 0
  
  # Replace Data With Interpolated Data
  replace_indeces = which(is.na(model_data[,i]))
  model_data[replace_indeces,i] = prediction[replace_indeces]
}
```

Below is some code to plot histograms of the data (the bin width will need to be changed based on the variables)
to compare the interpolated data to the original data
```{r}
# Create a Plot to Compare Data Sets before and After
distr_df = data.frame(Filled = model_data$percent_land_agricultural_2013, Missing = limited_data$percent_land_agricultural_2013) %>% 
  melt(value.name='y_i')

# Plot
ggplot(distr_df, aes(x = y_i, fill = variable)) +
  geom_histogram(position = "identity",binwidth = 2, alpha = .4) +
  labs(x = "y_i", y = "Frequency", 
       title = "Comparison") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 12))
```

# Output Data

```{r output_data}
# Save the Data for Use in the Next File
save(model_data, file="Intermediate_Data/03_model_data.Rdata")
```

